---
title: "tomato_egg_regression"
author: "Waveley Qiu"
date: "2023-04-10"
output: pdf_document
---

## Tomato Price Model Construction

$$
\begin{aligned}
\frac{y}{2} &= \beta_0 + \beta_1 x + \varepsilon
\\
\varepsilon &\sim N(0,\sigma^2),
\end{aligned}
$$

where $y$ is the price for a pound of tomatoes (as an average across cities in the US) and $x$ is time (in days from January 1, 1970). 

```{r, echo=TRUE}
tom_reg <- lm(value/2 ~ date, 
              data = tom_and_egg %>% 
                     filter(name == "tomato_price"))
```

```{r, echo=FALSE}
summary(tom_reg) %>%
  broom::tidy() %>%
  knitr::kable(col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"))
```

## Egg Price Model Construction

$$
\begin{aligned}
\log\Big(\frac{y}{3}\Big) &= \beta_0 + \beta_1 x + \varepsilon
\\
\varepsilon &\sim N(0,\sigma^2),
\end{aligned}
$$

where $y$ is the price for a dozen eggs (as an average across cities in the US)  and $x$ is time (in days from January 1, 1970). 

```{r egg_mod, echo=TRUE}
egg_reg <- lm(log(value/3) ~ date, 
              data = tom_and_egg %>% 
                     filter(name == "egg_price"))
```

```{r, echo=FALSE}
summary(egg_reg) %>%
  broom::tidy() %>%
  knitr::kable(col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"))
```

## Predicted Price of Ingredients on May 16, 2023

- Predicted price of a half pound of tomatoes on May 16, 2023:

$$
\begin{aligned}
\frac{\hat{y}}{2} &= \hat{\beta_0} + \hat{\beta_1}x
\\
&\approx -2.3379 + 0.0001565(\text{as.Date('May 16, 2023')})
\\
&\approx 0.7127
\end{aligned}
$$

- Predicted price of four eggs on May 16, 2023:

$$
\begin{aligned}
\log\Big(\frac{\hat{y}}{3}\Big) &= \hat{\beta_0} + \hat{\beta_1} x
\\
&\approx -11.2648 + 0.0006139(\text{as.Date('May 16, 2023')})
\\
&\approx 0.7019
\\
e^{0.7019} &\approx 2.017
\end{aligned}
$$
